The Weakly Nonlinear Magnetorotational Instability in a thin-gap Taylor-Couette flow

### 2017-02-17 paper/thingap_ref_response/thingap_mri_paper_ref_response_1.tex
new draft to respond to referee

\beq
\Psi = \partial_x \Psi = u_y = A = \partial_x B_y = 0.
\eeq

@@

@@

The fact that we find a Ginzburg-Landau equation for the amplitude is unsurprising: such an amplitude equation can be found in any system with Euclidean symmetry and a quadratic maximum in growth rate with respect to the wavenumber \citep{Hoyle:2006}. In this case, the Euclidean symmetry comes from axisymmetry in the $x$-$z$ plane, and the quadratic maximum is a consequence of the linear dispersion relation given in Appendix~\ref{app:dispersion}.  In Paper II, we show that the same symmetry occurs in the axisymmetric global geometry as well.

>>>>

\beq\label{eq:wallbcs}
\Psi = \partial_x \Psi = u_y = A = \partial_x B_y = 0.
\eeq

@@

\subsection{\textbf{Shearing box and ambipolar diffusion}}

Many studies of the MRI consider the instability in a shearing box, i.e. a wall-less local approximation that is meant to represent a small section of a disk. The shearing box obtains when Equations \ref{eqset1} - \ref{eqset4} are subjected to shear periodic radial boundary conditions rather than Equation \ref{eq:wallbcs}. The periodic nature of the shearing box allows us to decompose the fluid perturbations into Fourier modes proportional to $e^{i k_x x + i k_z z}$. This makes the shearing box MRI straightforward to treat analytically. However, as noted above, the fastest-growing linear MRI modes in the shearing box are also exact solutions of the nonlinear MRI equations -- that is, $J(\hat{\psi_0}, \hat{\psi_0}) = J(\hat{\psi_0}, \nabla^2 \hat{\psi_0}) = 0$ for $\hat{\psi_0} \propto e^{i k_x x + i k_z z}$. While this may be an appealing trait for analytic simplicity, it leads to the unphysical conclusion that the fastest growing modes will never nonlinearly interact. This `nonlinear property' will not be satisfied for two MRI modes with nonparallel wavenumbers, but with vertically periodic boundary conditions the most unstable mode has an axial vertical wavenumber. Thus a formal weakly nonlinear analysis cannot be conducted, as the most unstable mode will never nonlinearly interact with itself or its complex conjugate. Similarly, we cannot analytically examine interactions between MRI channel modes and damped eigenmodes belonging to other wave families. While the shearing box approximation allows

@@

The fact that we find a Ginzburg-Landau equation for the amplitude is unsurprising: such an amplitude equation can be found in any system with Euclidean symmetry and a quadratic maximum in growth rate with respect to the wavenumber \citep{Hoyle:2006}. In this case, the Euclidean symmetry comes from axisymmetry in the $x$-$z$ plane, and the quadratic maximum is a consequence of the linear dispersion relation given in Appendix~\ref{app:dispersion}. \textbf{The geometry of the problem considered here mimics an idealized experimental setup, where the local MRI equations are subject to wall-like radial boundary conditions.} In Paper II, we show that the same symmetry occurs in the axisymmetric global geometry as well. \textbf{How would the theory developed here change with the addition of vertical endcaps, as found in a real experiment?}%\textbf{Many studies of the MRI take place in the local shearing box, in which Equations \ref{eqset1} - \ref{eqset4} are solved subject to shear-periodic boundary conditions. This is meant }
